Deep Reinforcement Learning for Conservation Decisions

### 2021-06-08 appendices/appendix-A-definitions.Rmd
add pdfs
Moving exploration paragraph to the appendix

Exploration is a central component of any deep RL algorithm, and there are a variety of exploration algorithms that an agent can employ.
In the sections above, we procedurally mentioned that the agent must select actions during training.
Yet, how the agent selects these actions is very important because if an agent explores the state-action space poorly, the agent will learn a sub-optimal policy.
As mentioned in the manuscript, the dilemma in training is that it is difficult for the agent to know when to balance exploration of new actions and exploitation of past behavior.
For RL problems with small state-action spaces, e.g. space-action spaces that can be represented in a table, simple exploration algorithms will converge to the optimal policy, but with large or infinite state-action spaces, there are no convergence guarantees.
In the sections below, we will present some of the more commonly encountered exploration algorithms and then point to some more advanced methods.

@@

# Replay Buffers vs. Parallel Workers

Another critical algorithm design choice involves how an RL agent collects and samples past experiences.
A significant issue that arises from sampling past experiences is correlation of samples [@DQN, @schaul_prioritized_2016].
In the simple case that we have an RL agent that interacts with the environment for a few time steps and then uses this trajectory for its learning update, these samples, since they come from the same trajectory, will be strongly correlated.
Learning from correlated samples is undesirable since this can result in agents getting stuck in local optima, which will impede learning, but there are other reasons why learning from a single trajectory causes problems -- see @schaul_prioritized_2016 and @DQN for more insight here.
There are two commonly used solutions to this problem of correlated samples: Replay Buffers and Parallel Workers.
If the agent is off-policy, the agent can employ a replay buffer -- see @DQN for an example of a replay buffer-based agent.
With a replay buffer, the agent will save state-action-reward transitions in the replay buffer, and during learning updates, the agent will sample transitions from the replay buffer.
Since the replay buffer can be arbitrarily large and thus store transitions from numerous trajectories, sampling from the buffer will result in decorrelated samples.
On-policy agents, however, cannot use a replay buffer since these agents are not allowed to use samples from a policy different than the policy they are trying to improve -- note replay buffers are generally large enough to hold transitions from old policies.
The solution to de-correlate samples for on-policy agents is to use parallel workers, whereby multiple workers explore the environment according to the current policy.
Due to stochasticity that can come from exploration and the environment, the workers will go on different trajectories in state-action space.
The agent will collect samples across these workers in a learning step, perform its learning step and then repeat the process of collecting and sampling transitions with an updated policy.
The idea behind using parallel workers is that sampling over multiple trajectories de-correlates samples -- see @a3c on more details here.
In practice, using a large replay buffer, on the order of millions of transitions, or a modest number of parallel, around 4-16 for simple RL problems, is very important to achieve good agent performance for off-policy and on-policy algorithms respectively.

>>>>

Over the course of training, RL agents will engage in trial and error learning, whereby it will be necessary for the agent to explore new sequences of actions as well as exploit past high rewarding sequences.
In the sections above, we procedurally mentioned that the agent must select actions during training.
Yet, how the agent selects these actions is very important because if an agent explores the state-action space poorly, the agent will learn a sub-optimal policy.
Occasionally, the agent will need to propose exploratory actions that may result in higher rewards than previously seen.
Yet, the agent may also need to focus on past sequences of actions which may require marginal improvement to reach optimality.
The fundamental issue is that it is very difficult for the agent to know when it should engage in explorative or exploitative behavior.
This *exploration-exploitation dilemma* is an open research problem [@berger-tal_exploration-exploitation_2014].
For RL problems with small state-action spaces, e.g. space-action spaces that can be represented in a table, simple exploration algorithms will converge to the optimal policy, but with large or infinite state-action spaces, there are no convergence guarantees.

Current deep RL algorithms take a range of approaches towards this dilemma.
In the sections below, we will present some of the more commonly encountered exploration algorithms and then point to some more advanced methods.

@@

# On-Policy vs. Off-policy Evaluation

Before deciding how to explore the environment, RL algorithms must have a policy to use during training.
RL agents can either select actions from the same policy that they intend to use during evaluation or they can sample actions from a policy that is different.
The former is called on-policy learning; the agent learns to select actions with the same policy it will use in testing.
The latter is called off-policy learning; the agent uses a policy during training that is potentially different than the policy that the agent will evaluate.

A significant issue that arises when sampling actions and then using this experience to maximize the RL objective is correlation [@DQN, @schaul_prioritized_2016].
In the simple case that we have an RL agent that interacts with the environment for a few time steps and then uses this trajectory for its learning update, these samples, since they come from the same trajectory, will be strongly correlated[@DQN].
Learning from correlated samples is undesirable since this can result in agents getting stuck in local optima, which will impede learning, but there are other reasons why learning from a single trajectory causes problems -- see @schaul_prioritized_2016 and @DQN for more insight here.

On-policy and off-policy agents have different solutions to the correlated samples problem.
If the agent is off-policy, the agent can employ a replay buffer -- see @DQN for an example of a replay buffer-based agent.
With a replay buffer, the agent saves state-action-reward transitions in the replay buffer, and during learning updates, the agent samples transitions from the buffer.
Since the replay buffer can be arbitrarily large and thus store transitions from numerous trajectories, sampling from the buffer results in decorrelated samples.
On-policy agents, however, cannot use a replay buffer since these agents are not allowed to use samples from a policy different than the policy they are trying to improve -- note replay buffers are generally large enough to hold transitions from multiple policies.
The solution to de-correlate samples for on-policy agents is to use parallel workers, whereby multiple workers explore the environment according to the current policy.
Due to stochasticity that can come from exploration and the environment, the workers go on different trajectories in state-action space.
The agent will collect samples across these workers in a learning step, perform its learning step and then repeat the process of collecting and sampling transitions with an updated policy.
The idea behind using parallel workers is that sampling over multiple trajectories de-correlates samples -- see @a3c on more details here.
In practice, using a large replay buffer, on the order of millions of transitions, or a modest number of parallel, around 4-16 for simple RL problems, is very important to achieve good agent performance for off-policy and on-policy algorithms respectively.

### 2021-06-08 manuscript/manuscript.Rmd
Moving exploration paragraph to the appendix

---
title: Deep Reinforcement Learning for Conservation Decisions
authors:
  - name: Marcus Lapeyrolerie
    department: Department of Environmental Science, Policy, and Management
    affiliation: University of California, Berkeley
    location: Berkeley, California
  - name: Melissa Chapman
    department: Department of Environmental Science, Policy, and Management
    affiliation: University of California, Berkeley
    location: Berkeley, California
  - name: Kari Norman
    department: Department of Environmental Science, Policy, and Management
    affiliation: University of California, Berkeley
    location: Berkeley, California
  - name: Carl Boettiger
    department: Department of Environmental Science, Policy, and Management
    affiliation: University of California, Berkeley
    location: Berkeley, California
    email: cboettig@berkeley.edu
abstract: |
  Enter the text of your abstract here.
keywords:
  - blah
  - blee
  - bloo
  - these are optional and can be removed
bibliography: references.bib
header-includes:
  - \usepackage{amsmath}
  - \usepackage{amssymb}
output: 
  rticles::arxiv_article:
    keep_tex: true
  word_document: default
editor_options: 
  markdown: 
    wrap: sentence
---

@@

Neural networks become useful in RL when the environment has a large state-action space, which happens frequently with realistic decision-making problems. 
Classic RL algorithms, like policy iteration, value iteration and TD-learning, fail when the state-action space is large, e.g. when the possible states and actions can not be represented in a tractable table [@suttonbarto].
Yet, over the last 10 years, researchers have shown that neural network-based RL algorithms can perform well on previously unsolvable environments [@DQN, @alphaGo2016].
Neural networks have been widely useful in reinforcement learning because neural nets have the property of being general function approximators [@universalapproximators].
Whenever there is a need for an agent to approximate some function, say a policy function or value function, neural networks can be used in this capacity.
While there are other function approximators that can be used in RL, e.g. Gaussian processes [@grande14], neural networks have excelled in this role because of their ability to learn complex, non-linear, high dimensional functions and their ability to adapt given new information [@briefsurveydrl].
We will not discuss in detail how neural networks work [see @lecun_deep_2015], but the general process for how neural networks are used in RL is that at each learning step, the agent, will adjust the parameters in its neural network with which it approximates the policy function and/or value function, so that according to past experiences, the agent will be more likely to engage in highly rewarding behavior in the future. 
This same process occurs in value learning and actor-critic-based deep RL algorithms, except these agents are adjusting either a value network or both a value and policy network respectively

Over the course of training, RL agents will engage in trial and error learning, whereby it will be necessary for the agent to explore new sequences of actions as well as exploit past high rewarding sequences.
Occasionally, the agent will need to propose exploratory actions that may result in higher rewards than previously seen. 
The agent may also need to focus on past sequences of actions which may require marginal improvement to reach optimality.
It is very difficult for the agent to know when it should engage in explorative or exploitative behavior. 
This *exploration-exploitation dilemma* is an open research problem [@berger-tal_exploration-exploitation_2014]. 
Current deep RL algorithms take a range of approaches towards this (see  Appendix A).
The manner in which RL algorithms explore depend on whether the algorithms are *on-policy* or *off-policy*.
The distinction between on and off-policy evaluation comes from what policy the agent is using to select actions during training.
For off-policy algorithms, the agent uses a policy during training that can be different from the policy that the agent will use during evaluation; while in on-policy algorithms, the agent updates the same policy in training as the agent will use during evaluation.
Other points of divergence include whether the agent learns a deterministic or stochastic policy, and
whether the algorithm is formulated to work on a discrete, continuous or both a discrete and continuous state and action space.

@@

We provide two examples that illustrate the application and potential of deep RL to ecological and conservation problems, highlighting both the potential and the inherent challenges.

```{r knitr, include = FALSE}
library(tidyverse)
library(patchwork)
# This chunk controls plot aesthetics only, and can be omitted with no material change
knitr::opts_chunk$set(echo=FALSE, message = FALSE, warning = FALSE,
                      fig.width = 7, fig.height = 4, cache = FALSE)
ggplot2::theme_set(ggplot2::theme_bw())

@@

The first example focuses on the important but well-studied problem of setting harvest quotas in fisheries management.
Determining fishing quotas is both a critical ecological issue [@Worm2006; @Worm2009; @Costello2016], and also a textbook example that has long informed the management of renewable resources within fisheries and beyond [@Clark1990].

Given a population growth model that predicts the total biomass of a fish stock in the following year as a function of the current biomass, it is straight forward to determine what biomass corresponds to the maximum growth rate of the stock, or $B_{\textrm{MSY}}$, the biomass at Maximum Sustainable Yield (MSY) [@Schaefer1954].
When the population growth rate is stochastic, the problem is slightly harder to solve, as the harvest quota must constantly adjust to the ups and downs of stochastic growth, but it is still possible to show the optimal strategy merely seeks to maintain the stock at $B_{\textrm{MSY}}$, adjusted for any discounting of future yields [@Reed1979].

@@

For illustrative purposes, we consider the simplest version of the dynamic optimal harvest problem as outlined by [@Clark1973] (for the deterministic case) and [@Reed1979] (under stochastic recruitment).
The manager seeks to optimize the net present value (discounted cumulative catch) of a fishery, observing the stock size each year and setting an appropriate harvest quota in response.
In the classical approach, the best model of the fish population dynamics must first be estimated from data, potentially with posterior distributions over parameter estimates reflecting any uncertainty.
From this model, the optimal harvest policy -- that is, the function which returns the optimal quota for each possible observed stock size -- can be determined either by analytic [@Reed1979] or numerical [@Marescot2013] methods, depending on the complexity of the model.
In constrast, a model-free deep RL algorithm makes no assumption about the precise functional form or parameter values underlying the dynamics -- it is in principle agnostic to the details of the simulation.

We illustrate the deep RL approach using the algorithm known as Twin Delayed Deep Deterministic Policy Gradient or more simply, TD3 [@TD3].
TD3 is an extension of Deep Deterministic Policy Gradient [DDPG, @DDPG], which itself extended the original Deep Q-Network [DQN, @DQN] algorithm to continuous action space.
A step-by-step walk-through for training both agents on this environment is provided in the Appendix.
We compare the resulting management, policy, and reward under either RL agent to that achieved by the optimal management solution [Fig 2].
Despite having no knowledge of the underlying model, the RL agent learns enough to achieve nearly optimal performance.

@@

The cumulative reward (utility) realized across 100 stochastic replicates is indistinguishable from that of the optimal policy [Fig 2].
Nevertheless, comparing the mean state over replicate simulations reveals some differences in the RL strategy, wherein the stock is maintained at a slightly higher-than-optimal biomass.
Because our state space and action space are sufficiently low-dimensional in this example, we are also able to visualize the policy function directly, and compare to the optimal policy [Fig 2].
This confirms that quotas tend to be slightly lower than optimal, most notably at larger stock sizes.
These features highlight a common challenge in the design and training of RL algorithms.
RL cares only about improving the realized cumulative reward, and may sometimes achieve this in unexpected ways.
Because these simulations rarely reach stock sizes at or above carrying capacity, these larger stock sizes show a greater deviation from the optimal policy than we observe at more frequently visited lower stock sizes.
Training these agents in a variety of alternative contexts can improve there ability to generalize to other scenarios.

How would an RL agent be applied to empirical data?
In principle, this is straight forward.
Once our agent has been trained on appropriate simulation environments, it is able to propose a quota given an observation of the stock after appropriate transformation of variables.
To illustrate this, we examine the quota that would be recommended by our newly trained RL agent, above, against historical harvest levels of Argentine hake based on stock assessments from 1986 - 2014 [@ram, see Appendix D].
Hake stocks showed a marked decline throughout this period, while harvests deceased only in proportion [Fig 3].
In contrast, our RL agent would have recommended significantly lower quotas over most of the same interval, including the closure of the fishery as stocks were sufficiently depleted.
While we have no way of knowing for sure if the RL quotas would have led to recovery, let alone an optimal harvest rate, the contrast between those recommended quotas and the historical catch is notable.

@@

This approach is not as different from conventional strategies at it may seem.
In a conventional approach, ecological models are first estimated from empirical data, (stock assessments in the fisheries case).
Quotas can then be set based directly on these model estimates, or by comparing alternative candidate "harvest control rules" (the policy function) against model-based simulations of stock dynamics.
This latter approach, known in fisheries as Management Strategy Evaluation [MSE; @Punt2016] is already very closely analogous to the RL process.
Instead of researchers evaluating a handful of control rules, the RL agent takes on the task of proposing and evaluating the millions of possible control rules represented by its neural network.

@@

Our second example focuses on a case for which we do not already have an existing, provably optimal policy to compare against.
We consider the generic problem of an ecosystem facing slowly deteriorating environmental conditions which move the dynamics closer towards a tipping point [Fig 2].
This model of a critical transition has been posited widely in ecological systems, from the simple consumer-resource model of [@May1977] on which our dynamics are based, to microbial dynamics [@Dai2012] lake ecosystem communities [@Carpenter2011] to planetary ecosystems [@Barnosky2012].
On top of these ecological dynamics we introduce an explicit ecosystem service model quantifying the value of more desirable 'high' state relative to the 'low' state.
For simplicity, we assume a proportional benefit $b$ associated with the ecosystem state $X(t)$.
Thus when the ecosystem is near the 'high' equilibrium $\hat X_H$, the corresponding ecosystem benefit $b \hat X_H$ is higher than at the low equilibrium, $b x_L$, consistent with the intuitive description of ecosystem tipping points [@Barnosky2012].

We also enumerate the possible actions which a manager may take in response to environmental degradation.
In the absence of any management response, we assume the environment deteriorates at a fixed rate $\alpha$, which can be thought of as the incremental increase in global mean temperature or similar anthropogenic forcing term.
Management can respond to slow or even reverse this trend by choosing an opposing action $A_t$.
We assume that large actions are proportionally more costly than small actions, consistent with the expectation of diminishing returns: taking the cost associated with an action $A_t$ as equal to $c A_t^2$.
Many alterations of these basic assumptions are also possible: our `gym_conservation` implements a range of different scenarios with user-configurable settings.
In each case, the manager observes the current state of the system each year and must then select the policy response that year.

@@

A close examination of the trajectories of individual simulations which cross the tipping point under either management strategy [see appendix B] further highlights the difference between these two approaches.
Under the steady-state strategy, the system remains poised too close to the tipping point: stochastic noise eventually drives most replicates across the threshold, where the steady-state strategy is too weak to bring them back once they collapse.
As replicate after replicate stochastically crashes, the mean state and mean reward bend increasingly downwards.
In contrast, the RL agent edges the system slightly farther away from the tipping point, decreasing but not eliminating the chance of a chance transition.
In the few replicates that experience a critical transition anyway, the RL agent usually responds with sufficient commitment to ensure their recovery [Appendix B].
Only 3 out of 100 replicates degrade far enough for the RL agent to give up the high cost of trying to rescue them.
The RL agent's use of a more dynamic strategy out-preforms the steady-state strategy.
Numerous kinks visible in the RL policy function also suggest that this solution is not yet optimal.
Further tuning of hyper-parameters, increased training, alterations or alternatives to the training algorithm would likely be able to further improve upon this performance.
Nevertheless, such quirks are likely to be common features of RL solutions -- long as they have minimal impact on realized rewards.

@@

>>>>

---
title: Deep Reinforcement Learning for Conservation Decisions
authors:
  - name: Marcus Lapeyrolerie
    department: Department of Environmental Science, Policy, and Management
    affiliation: University of California, Berkeley
    location: Berkeley, California
  - name: Melissa Chapman
    department: Department of Environmental Science, Policy, and Management
    affiliation: University of California, Berkeley
    location: Berkeley, California
  - name: Kari Norman
    department: Department of Environmental Science, Policy, and Management
    affiliation: University of California, Berkeley
    location: Berkeley, California
  - name: Carl Boettiger
    department: Department of Environmental Science, Policy, and Management
    affiliation: University of California, Berkeley
    location: Berkeley, California
    email: cboettig@berkeley.edu
abstract: |
  Enter the text of your abstract here.
keywords:
  - blah
  - blee
  - bloo
  - these are optional and can be removed
bibliography: references.bib
header-includes:
  - \usepackage{amsmath}
  - \usepackage{amssymb}
output: 
  rticles::arxiv_article:
    keep_tex: true
editor_options: 
  markdown: 
    wrap: sentence
---

@@

Neural networks become useful in RL when the environment has a large state-action space, which happens frequently with realistic decision-making problems. 
Classic RL algorithms, like policy iteration, value iteration and TD-learning, fail when the state-action space is large, e.g. when the possible states and actions can not be represented in a tractable table [@suttonbarto].
Yet, over the last 10 years, researchers have shown that neural network-based RL algorithms can perform well on previously unsolvable environments [@DQN, @alphaGo2016].
Neural networks have been widely useful in reinforcement learning because neural nets have the property of being general function approximators [@universalapproximators].
Whenever there is a need for an agent to approximate some function, say a policy function or value function, neural networks can be used in this capacity.
While there are other function approximators that can be used in RL, e.g. Gaussian processes [@grande14], neural networks have excelled in this role because of their ability to learn complex, non-linear, high dimensional functions and their ability to adapt given new information [@briefsurveydrl].
We will not discuss in detail how neural networks work -- see [@lecun_deep_2015] for further background --, but the general process for how neural networks are used in RL is that at each learning step, the agent, say a policy gradient-based agent, will adjust the parameters in its policy-approximating neural network so that according to past experiences, the agent will be more likely to engage in highly rewarding behavior in the future. 
This same process occurs in value learning and actor-critic-based deep RL algorithms, except these agents are adjusting either a value network or both a value and policy network respectively

On top of the aforementioned distinctions of deep RL algorithms, there are more algorithms design choices that are made. 
A very significant consideration is deciding how the agent will explore the environment.
Another is whether the algorithm can work on a discrete, continuous or both a discrete and continuous state and action space.
In Appendix A, we provide more detail on the important components of RL algorithms.

@@

We provide three examples that illustrate the application and potential of deep RL to ecological and conservation problems, highlighting both the potential and the inherent challenges.

```{r knitr, include = FALSE}
library(tidyverse)
library(patchwork)
# This chunk controls plot aesthetics only, and can be omitted with no material change
knitr::opts_chunk$set(echo=TRUE, message = FALSE, warning = FALSE,
                      fig.width = 7, fig.height = 4, cache = FALSE)
ggplot2::theme_set(ggplot2::theme_bw())

@@

The first example focuses on the important but well-studied problem of setting harvest quotas in fisheries management.
Determining fishing quotas is both a critical ecological issue [@Worm2006; @Worm2009; @Costello2016], and also a textbook example that has long informed the management of renewable resources within fisheries and beyond [@Clark1990].
Given a population growth model that predicts the total biomass of a fish stock in the following year as a function of the current biomass, it is straight forward to determine what biomass corresponds to the maximum growth rate of the stock, or $B_{\textrm{MSY}}$, the biomass at Maximum Sustainable Yield (MSY) [@Schaefer1954].
When the population growth rate is stochastic, the problem is slightly harder to solve, as the harvest quota must constantly adjust to the ups and downs of stochastic growth, but it is still possible to show the optimal strategy merely seeks to maintain the stock at $B_{\textrm{MSY}}$, adjusted for any discounting of future yields [@Reed1979].

@@

For illustrative purposes, we consider the simplest version of the dynamic optimal harvest problem as outlined by [@Clark1973] (for the deterministic case) and [@Reed1979] (under stochastic recruitment).
In short, the manager seeks to optimize the net present value (discounted cumulative catch) of a fishery, observing the stock size each year and setting an appropriate harvest quota in response.
In the classical approach, the best model of the fish population dynamics must first be estimated from data, potentially with posterior distributions over parameter estimates reflecting any uncertainty.
From this model, the optimal harvest policy -- that is, the function which returns the optimal quota for each possible observed stock size -- can be determined either by analytic [@Reed1979] or numerical [@Marescot2013] methods, depending on the complexity of the model.
However, unlike an analytic or SDP solution that is derived directly from the estimated model, a model-free deep RL algorithm makes no assumption about the precise functional form or parameter values underlying the dynamics -- it is in principle agnostic to the details of the simulation.

Here, we apply a version of the deep RL algorithm known as Twin Delayed Deep Deterministic Policy Gradient or more simply, TD3 [@TD3].
A step-by-step walk-through for training both agents on this environment is provided in the Appendix.
We compare the resulting management, policy, and reward under either RL agent to that achieved by the optimal management solution [Fig 2].
Despite having no knowledge of the underlying model and clear differences in the resulting strategies, both RL agents learn enough to achieve nearly optimal performance.

@@

These features highlight a common challenge in the design and training of RL algorithms.
RL cares only about improving the realized cumulative reward, and may sometimes achieve this in unexpected ways.
Because these simulations rarely reach stock sizes at or above carrying capacity, these larger stock sizes show a greater deviation from the optimal policy than we observe at lower stock sizes.
As intuition would suggest, it is possible to devise alternative training schemes which may do a better job sampling the simulation dynamics at high stock levels.
Training these agents in a variety of alternative contexts can improve there ability to generalize to other scenarios.

<!-- 
Add text about empirical example
-->

@@

@@

Our second example focuses on a case for which we do not already have an existing, provably optimal policy to compare against.
We consider the generic problem of an ecosystem facing slowly deteriorating environmental conditions which move the dynamics closer towards a tipping point [Fig 2].
This model of a critical transition has been posited widely in ecological systems, from the simple consumer-resource model of [@May1977] on which our dynamics are based, to microbial dynamics [@Dai2012] lake ecosystem communities [@Carpenter2011] to planetary ecosystems [@Barnosky2012].
On top of these ecological dynamics we introduce an explicit ecosystem service model quantifying the value of more desirable 'high' state relative to the 'low' state.
For simplicity, we assume a proportional benefit $b$ associated with the ecosystem state $X(t)$.
Thus when the ecosystem is near the 'high' equilibrium $\hat X_H$, the corresponding ecosystem benefit $b \hat X_H$ is higher than at the low equilibrium, $b x_L$, consistent with the intuitive description of ecosystem tipping points [@Barnosky2012].
We also enumerate the possible actions which a manager may take in response to environmental degradation.
In the absence of any management response, we assume the environment deteriorates at a fixed rate $\alpha$, which can be thought of as the incremental increase in global mean temperature or similar anthropogenic forcing term.
Management can respond to slow or even reverse this trend by choosing an opposing action $A_t$.
We assume that large actions are proportionally more costly than small actions, consistent with the expectation of diminishing returns: taking the cost associated with an action $A_t$ as equal to $c A_t^2$.
Many alterations of these basic assumptions are also possible: our `gym_conservation` implements a range of different scenarios with user-configurable settings.
In each case, the manager observes the current state of the system each year and must then select the policy response that year.

@@

A close examination of the trajectories of individual simulations which cross the tipping point under either management strategy (see appendix B, fig S1, S2) further highlights the difference between these two approaches.
Under the steady-state strategy, the system remains poised too close to the tipping point: stochastic noise eventually drives most replicates across the threshold, where the steady-state strategy is too weak to bring them back once they collapse.
As replicate after replicate stochastically crashes, the mean state and mean reward bend increasingly downwards.
In contrast, the RL agent edges the system slightly farther away from the tipping point, decreasing but not eliminating the chance of a chance transition.
In the few replicates that experience a critical transition anyway, the RL agent usually responds with sufficient commitment to ensure their recovery (Fig S2).
Only 3 out of 100 replicates degrade far enough for the RL agent to give up the high cost of trying to rescue them.

Such examination makes it clear how the RL agent's use of a more dynamic strategy out-preforms the steady-state strategy.
The various kinks in the RL policy function also suggest that this solution is nevertheless surely not the optimal solution -- such features are typical of RL-based policies, where such spandral-like features from early experimentation may become baked into the long-term strategy if adjusting them has little improvement on the final reward. <!-- ML: the parenthetical here I understand but I feel a little ehh on the word choice. I also don't know how to rephrase it -->
Further tuning of hyper-parameters, increased training, alterations or alternatives to the training algorithm would likely be able to further improve upon this performance.

@@

<!-- Something about the way forward, point to appendix resources etc -->

<!-- Discussion really doesn't discuss/draw conclusions from the fisheries and tipping points examples! Do we need to or maybe not since it's not about the science-->

<!-- ML: I think talking about the examples in a very passing manner is fine as the point is primarily to show that RL can work. To conclude, I feel like we could portray the vision of open ai gym environments building a CS-ecology community.-->

<!-- 
The pitfalls of RL-based solutions to global environmental change problems are not only technical, but include issues of ethics and power, especially when the algorithms or data are proprietary.
We discuss not only technical issues that can cause RL solutions to fail in surprising ways, but also potential issues raised by the adoption of more algorithmic approaches to conservation generally.
We emphasize how an open, transparent and reproducible approach can help mitigate some (but not all) of these concerns, and facilitate a more effective interface between teams of researchers from both ecological and computer sciences.
If algorithms are to play a meaningful role in the allocation and exploitation of natural resources from fish to forests, good intentions [@Joppa] may not be sufficient to ensure equitable outcomes.
Ecologists and computer scientists will also need input from experts in governance and ethics to ensure that applications of AI to ecological challenges are fair and just.
